Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-20 paper_jasa.tex
plot 7 updated
plot 3 updated

%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
%\begin{thm}\label{thm:power}
% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
%\end{thm}
% 
%{ \em proof}
%
%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
%\begin{eqnarray*}
%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
%
%Thus expected power in equation \ref{power_exp} be
%\begin{eqnarray*}
% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
%  &=& E(1-p_D)^{m-1}   \\
%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%
\begin{figure}[hbtp]
   \centering
       \scalebox{.95}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
       \label{fig:power_expected}
\end{figure}

@@

Model \ref{mixed} with effect as fixed covariate and subject as random effect is fitted to the data to estimate the subject specific power curves shown in figure \ref{fig:power_mixed_subject}. Subject specific power is quite different for three experiments. It appears that scatter plot serves better job compared to boxplot as a visual test statistic. But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.

>>>>

%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
%\begin{thm}\label{thm:power}
% Equation \ref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
%\end{thm}
% 
%{ \em proof}
%
%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
%\begin{eqnarray*}
%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
%
%Thus expected power in equation \ref{power_exp} be
%\begin{eqnarray*}
% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
%  &=& E(1-p_D)^{m-1}   \\
%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%
\begin{figure}[hbtp]
   \centering
       \scalebox{.6}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
       \label{fig:power_expected}
\end{figure}

@@

In all the three experiments, each subject saw a lineup only once. Assuming that there is no random effect for lineup, we set  $\tau_{\ell  i}=\tau_i $ in model \ref{mixed} and $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated. This gives us the estimate of subject specific probability of successful evaluation of a lineup ($\hat p_i$) from model \ref{mixed}. Parameter estimates of the model is shown in table \ref{tbl:model_par}.  The estimated subject specific power curves are shown in figure \ref{fig:power_mixed_subject}. The offset is set to $log(0.05/0.95)$ so that Type-I error is fixed to be 0.05 when effect = 0. Subject specific power is quite different for three experiments. It appears that scatter plot serves better job compared to boxplot as a visual test statistic. But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.

\begin{table}[hbtp]
\caption{Parameter estimates of model \ref{mixed}.}
\begin{center}
\begin{tabular}{ccc}
  \hline
 &  \multicolumn{2}{c} {Estimates}  \\
 \cline{2-3}
 Experiment & Effect & inter subject correlation\\
  \hline
  1 &  &  \\ 
  2 &  &  \\ 
  3 &  & \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:model_par}
\end{table}
